---
title: "Park area provision: Metric comparisons"
author: "Song, Xiao Ping"
date: "`r Sys.Date()`"
opengraph:
  image: 
    src: "man/figures/logo.png"
output: rmarkdown::html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r eval = TRUE, include = FALSE}
library(home2park)
```

```{r include = FALSE, eval = FALSE}
devtools::load_all() # to knit manually w latest changes (not built/installed yet)
# vignette will be checked in R-CMD check & package build (should use library(home2park))
```

```{r load data, include = FALSE}
data(buildings_pop_sgp)
buildings_pop_sgp <- sf::st_transform(buildings_pop_sgp, sf::st_crs(32648)) # transform to projected crs

data(parks_sgp)
parks_sgp <- sf::st_transform(parks_sgp, sf::st_crs(32648)) # transform to projected crs

data(pop_sgp)
pop_sgp <- sf::st_transform(pop_sgp, sf::st_crs(32648)) # transform to projected crs
```

```{r calc distance matrix and supply of park area, message = FALSE, warning = FALSE, include = FALSE}

# convert buildings to points (centroids), then calculate distances to every park
m_dist <- buildings_pop_sgp %>%
  sf::st_centroid() %>%
  sf::st_distance(parks_sgp) %>%
    units::set_units(NULL)


# new column for the supply of park area
buildings_pop_sgp$area_supply <- recre_supply(park_attribute = parks_sgp$area, 
                                              dist_matrix = m_dist, 
                                              c = 0.302) # e.g. from Tu et al. (2020)

# supply to all residents per building
buildings_pop_sgp$area_supplytopop <- buildings_pop_sgp$area_supply * buildings_pop_sgp$popcount



# scale the supply variables if necessary
buildings_pop_sgp <- buildings_pop_sgp %>%
    dplyr::mutate(across(.cols = c(area_supply, area_supplytopop), 
                         ~(scale(., center = FALSE))))

rm(m_dist)
```

```{r include = FALSE}
# calculate for year 2020
pop_2020 <- pop_sgp %>%
  dplyr::filter(year == 2020)
```

```{r summarise supply of park area per census block, message = FALSE, warning = FALSE, include = FALSE}

# append subzone info to buildings if building intersects with subzone  
buildings_subzones <- buildings_pop_sgp  %>%
  sf::st_join(pop_2020,
              join = sf::st_intersects,
              left = TRUE) %>%
  sf::st_set_geometry(NULL)


# summarise total/median/average supply value for all buildings within subzone
buildings_subzones <- buildings_subzones %>%
  dplyr::group_by(subzone_n) %>%
  dplyr::summarise(across(.cols = c(area_supply, area_supplytopop), 
                          .fns = sum, .names = "{.col}_sum"),
                   across(.cols = c(area_supply, area_supplytopop), 
                          .fns = median, .names = "{.col}_median"),
                   across(.cols = c(area_supply, area_supplytopop), 
                          .fns = mean, .names = "{.col}_mean")) 
  

# join information to pop_2020, calculate supply per capita
pop_2020 <- pop_2020 %>%
  dplyr::left_join(buildings_subzones) %>%
  dplyr::mutate(area_supplyperpop = area_supplytopop_sum / pop_count) %>%
  dplyr::mutate(area_supplyperpop = ifelse(is.infinite(area_supplyperpop), NA, area_supplyperpop)) # make infinite NA

```

```{r include = FALSE}
sf::st_agr(pop_2020) = "constant"
sf::st_agr(parks_sgp) = "constant"
# https://github.com/r-spatial/sf/issues/406
```


```{r conventional park provision metric - per capita park area, message = FALSE, warning = FALSE, include = FALSE}

subzones_parks <- sf::st_intersection(pop_2020, parks_sgp) # subset census block polygons to those that intersect parks, append park info
subzones_parks$parkarea_m2 <- sf::st_area(subzones_parks) # calc area of each polygon


# calculate total park area per census block
subzones_parks <- subzones_parks %>%
  dplyr::group_by(subzone_n) %>%
  dplyr::summarise(parkarea_m2 = as.numeric(sum(parkarea_m2))) %>%
  sf::st_drop_geometry()


# join information to pop_2020, calculate park area per capita
pop_2020 <- pop_2020 %>%
  dplyr::left_join(subzones_parks) %>%
  dplyr::mutate(parkarea_m2 = ifelse(is.na(parkarea_m2), 0, parkarea_m2)) %>% # make NA 0
  dplyr::mutate(parkperpop_m2 = parkarea_m2 / pop_count) %>%
  dplyr::mutate(parkperpop_m2 = ifelse(is.infinite(parkperpop_m2), # make infinite NA
                                NA, parkperpop_m2)) %>%
  dplyr::mutate(parkperpop_m2 = ifelse(is.nan(parkperpop_m2), # make NaN NA
                                NA, parkperpop_m2))

```

```{r include = FALSE}
rm(buildings_subzones, subzones_parks)
```

<br>

The following interactive map provides a visual comparison between the metrics for the provision of park area that were calculated in the '[Get started](https://ecological-cities.github.io/home2park/articles/home2park.html)' vignette.

```{r  plot park metrics on map, echo = FALSE, message = FALSE, warning = FALSE, dpi = 300, fig.height = 2.0, fig.width = 2.35, fig.cap=paste0("**Map: Supply of park area in Singapore based on OSM data (2020).** Only a randomly sampled subset of buildings are shown. All color palettes are binned according to quantile values.")}

# convert buildings to points (centroids) for plotting
buildings_pop_sgp <- buildings_pop_sgp %>% 
  sf::st_centroid() %>%
  dplyr::mutate(across(everything(), as.vector)) # remove attributes from columns for plotting



tmap::tmap_mode("view")

tm <- 
  tmap::tm_basemap(c("Esri.WorldGrayCanvas", "CartoDB.DarkMatter", "OpenStreetMap")) +

  # census blocks
  tmap::tm_shape(pop_2020) +
    tmap::tm_polygons(title = "Park area per capita (Ha/1000 residents)",
                      group = "Census blocks: Per capita park area (conventional)",
                      col = "parkperpop_m2",
                      palette = "Greens", alpha = 0.7,
                      style = "quantile",
                      border.col = "white", border.alpha = 0.5, lwd = 1) +
  tmap::tm_shape(pop_2020) +
    tmap::tm_polygons(title = "Supply of park area per capita (scaled)",
                      group = "Census blocks: Per capita supply of park area",
                      col = "area_supplyperpop", 
                      palette = "Greens", alpha = 0.7,
                      style = "quantile",
                      border.col = "white", border.alpha = 0.5, lwd = 1) +
          
  # buildings
  tmap::tm_shape(dplyr::slice_sample(buildings_pop_sgp, # subset random sample
                                     n = nrow(buildings_pop_sgp)/12)) +
    tmap::tm_dots(title = "Supply of park area (scaled)",
                  group = "Buildings: Supply of park area to residents",
                  col = "area_supplytopop", border.col = "transparent",
                  palette = viridis::viridis(5),
                  style = "quantile",
                  size = 0.01,
                  alpha = 0.7) +
  
  # parks  
  tmap::tm_shape(parks_sgp %>% dplyr::select(id, name, area)) +
    tmap::tm_polygons(group = "Parks",
                      col = "#33a02c",
                      alpha = 0.6,
                      border.col = "grey50",  border.alpha = 0.5)
  
  
# Pipe the tmap object into tmap_leaflet() to create a leaflet widget,
# so that we can use leaflet::hideGroup().
tm %>% 
  tmap::tmap_leaflet() %>%
  leaflet::hideGroup("Census blocks: Per capita park area (conventional)") %>%
  leaflet::hideGroup("Census blocks: Per capita supply of park area")
```


```{r include = FALSE}
rm(tm, buildings_pop_sgp, parks_sgp, pop_sgp, pop_2020)
```


<br>

## Data sources

- Singapore census data from the [Department of Statistics Singapore](https://www.singstat.gov.sg/find-data/search-by-theme/population/geographic-distribution/latest-data). Released under the terms of the [Singapore Open Data Licence version 1.0](https://data.gov.sg/open-data-licence).

- Singapore census block (subzone) polygons from the [Singapore Master Plan Subzones](https://data.gov.sg/dataset/master-plan-2019-subzone-boundary-no-sea). Released under the terms of the [Singapore Open Data Licence version 1.0](https://data.gov.sg/open-data-licence).

- Singapore Master Plan Land Use Zones for the years [2014](https://data.gov.sg/dataset/master-plan-2014-land-use) and [2019](https://data.gov.sg/dataset/master-plan-2019-land-use-layer). Released under the terms of the [Singapore Open Data License](https://data.gov.sg/open-data-licence).

- Building polygons derived from map data [copyrighted](https://www.openstreetmap.org/copyright) OpenStreetMap contributors and available from https://www.openstreetmap.org. Released under the terms of the [ODbL License](https://opendatacommons.org/licenses/odbl/summary/).

- Park polygons and summarised attributes (trails, playgrounds) derived from map data [copyrighted](https://www.openstreetmap.org/copyright) OpenStreetMap contributors and available from https://www.openstreetmap.org. Released under the terms of the [ODbL License](https://opendatacommons.org/licenses/odbl/summary/).


