---
title: "Park area provision: Metric comparisons"
author: "Song, Xiao Ping"
date: "`r Sys.Date()`"
opengraph:
  image: 
    src: "man/figures/logo.png"
output: rmarkdown::html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r eval = TRUE, include = FALSE}
library(home2park)
```

```{r include = FALSE, eval = FALSE}
devtools::load_all() # to knit manually w latest changes (not built/installed yet)
# vignette will be checked in R-CMD check & package build (should use library(home2park))
```

```{r load data, include = FALSE}
data(buildings_pop_sgp)
buildings_pop_sgp <- sf::st_transform(buildings_pop_sgp, sf::st_crs(32648)) # transform to projected crs

data(parks_sgp)
parks_sgp <- sf::st_transform(parks_sgp, sf::st_crs(32648)) # transform to projected crs

data(pop_sgp)
pop_sgp <- sf::st_transform(pop_sgp, sf::st_crs(32648)) # transform to projected crs
```

```{r calc distance matrix and supply of park area, message = FALSE, warning = FALSE, include = FALSE}

# convert buildings to points (centroids), then calculate distances to every park
m_dist <- buildings_pop_sgp %>%
  sf::st_centroid() %>%
  sf::st_distance(parks_sgp) %>% # euclidean distance
    units::set_units(NULL)

m_dist <- m_dist / 1000 # convert distances to km


# new column for the supply of park area
buildings_pop_sgp$area_supply <- recre_supply(park_attribute = parks_sgp$area, 
                                              dist_matrix = m_dist, 
                                              c = 0.302) # e.g. from Tu et al. (2020)

# supply to all residents per building
buildings_pop_sgp$area_supplytopop <- buildings_pop_sgp$area_supply * buildings_pop_sgp$popcount

rm(m_dist)
```

```{r include = FALSE}
# calculate for year 2020
pop_2020 <- pop_sgp %>%
  dplyr::filter(year == 2020)
```

```{r summarise supply of park area per census unit, message = FALSE, warning = FALSE, include = FALSE}

# append subzone info to buildings if building intersects with subzone  
buildings_subzones <- buildings_pop_sgp  %>%
  sf::st_join(pop_2020,
              join = sf::st_intersects,
              left = TRUE) %>%
  sf::st_set_geometry(NULL)


# summarise total/median/average supply value for all buildings within subzone
buildings_subzones <- buildings_subzones %>%
  dplyr::group_by(subzone_n) %>%
  dplyr::summarise(across(.cols = c(area_supply, area_supplytopop), 
                          .fns = sum, .names = "{.col}_sum"),
                   across(.cols = c(area_supply, area_supplytopop), 
                          .fns = median, .names = "{.col}_median"),
                   across(.cols = c(area_supply, area_supplytopop), 
                          .fns = mean, .names = "{.col}_mean")) 
  

# join information to pop_2020, calculate supply per capita
pop_2020 <- pop_2020 %>%
  dplyr::left_join(buildings_subzones) %>%
  dplyr::mutate(area_supplyperpop = area_supplytopop_sum / pop_count * 1e-6) %>% # convert to km2
  dplyr::mutate(area_supplyperpop = ifelse(is.infinite(area_supplyperpop), NA, area_supplyperpop)) # make infinite NA

```

```{r include = FALSE}
sf::st_agr(pop_2020) = "constant"
sf::st_agr(parks_sgp) = "constant"
# https://github.com/r-spatial/sf/issues/406
```


```{r conventional park provision metric - per capita park area, message = FALSE, warning = FALSE, include = FALSE}

subzones_parks <- sf::st_intersection(pop_2020, parks_sgp) # subset census unit polygons to those that intersect parks, append park info
subzones_parks$parkarea_m2 <- sf::st_area(subzones_parks) # calc area of each polygon


# calculate total park area per census unit
subzones_parks <- subzones_parks %>%
  dplyr::group_by(subzone_n) %>%
  dplyr::summarise(parkarea_m2 = as.numeric(sum(parkarea_m2))) %>%
  sf::st_drop_geometry()


# join information to pop_2020, calculate park area per capita
pop_2020 <- pop_2020 %>%
  dplyr::left_join(subzones_parks) %>%
  dplyr::mutate(parkarea_m2 = ifelse(is.na(parkarea_m2), 0, parkarea_m2)) %>% # make NA 0
  dplyr::mutate(parkperpop_m2 = parkarea_m2 / pop_count * 1e-6) %>% # convert to km2
  dplyr::mutate(parkperpop_m2 = ifelse(is.infinite(parkperpop_m2), # make infinite NA
                                NA, parkperpop_m2)) %>%
  dplyr::mutate(parkperpop_m2 = ifelse(is.nan(parkperpop_m2), # make NaN NA
                                NA, parkperpop_m2))

```

```{r include = FALSE}
rm(buildings_subzones, subzones_parks)
```

<br>

The following interactive map provides a visual comparison between metrics for park area provision calculated in the '[Get started](https://ecological-cities.github.io/home2park/articles/home2park.html)' vignette. Toggle the map layers to view the per capita provision of park area summarised per region (subzone).

```{r  plot park metrics on map, echo = FALSE, message = FALSE, warning = FALSE, dpi = 300, fig.height = 2.0, fig.width = 2.35, fig.cap=paste0("**Map: Supply of park area in Singapore based on OSM data (2020).** Each building is denoted as a point (a random subset is shown). All color palettes are binned according to quantile values.")}

# convert buildings to points (centroids) for plotting
buildings_pop_sgp <- buildings_pop_sgp %>% 
  sf::st_centroid() %>%
  dplyr::mutate(across(everything(), as.vector)) # remove attributes from columns for plotting


# for area related vars, convert m2 to km2
buildings_pop_sgp <- buildings_pop_sgp %>% 
  dplyr::mutate(across(.cols = contains("area"), 
              .fns = function(x) x*1e-6))


# random sampling of buildings
# set.seed(123) 
# # some processing to get proper range for color scale later
# # get max values per building (for particular round)
# buildings_max <- buildings_pop_sgp %>%
#   tidyr::pivot_longer(cols = c("area_supply", "area_supplytopop"),
#                       names_to = "supply", 
#                       values_to = "value") %>%
#      dplyr::group_by(supply) %>%
#      dplyr::slice(which.max(value)) %>%
#   tidyr::pivot_wider(names_from = "supply", values_from = "value")
# 
# # subset random sample (already done with example dataset)
# buildings_pop_sgp <-
#   dplyr::slice_sample(buildings_pop_sgp,
#                       n = nrow(buildings_pop_sgp)/20)
# 
# # append max buildings to the random subset 
# buildings_pop_sgp <- buildings_pop_sgp %>%
#   dplyr::bind_rows(buildings_max)





tmap::tmap_mode("view")
tmap::tmap_options(check.and.fix = TRUE)

tm <- 
  tmap::tm_basemap(c("Esri.WorldGrayCanvas", "CartoDB.DarkMatter", "OpenStreetMap")) +

  # parks  
  tmap::tm_shape(parks_sgp %>% dplyr::select(id, name, area)) +
    tmap::tm_polygons(group = "Parks",
                      col = "#33a02c",
                      alpha = 0.6,
                      border.col = "grey50",  border.alpha = 0.5) +

  # buildings
  tmap::tm_shape(buildings_pop_sgp) +
    tmap::tm_dots(title = "Buildings: Supply of park area (km<sup>2</sup>)",
                  group = "Buildings: Supply of park area",
                  col = "area_supply", border.col = "transparent",
                  palette = viridis::viridis(5),
                  style = "quantile",
                  size = 0.01,
                  alpha = 0.7,
                  interactive = FALSE, # some glitch in values of hover text/popup
                  popup.vars = NULL,
                  showNA = FALSE) +
  
  tmap::tm_shape(buildings_pop_sgp) +
    tmap::tm_dots(title = "All building residents: Supply of park area (km<sup>2</sup>)",
                  group = "Buildings: Supply of park area to residents",
                  col = "area_supplytopop", border.col = "transparent",
                  palette = viridis::viridis(5),
                  style = "quantile",
                  size = 0.01,
                  alpha = 0.7,
                  interactive = FALSE, # some glitch in values of hover text/popup
                  popup.vars = NULL,
                  showNA = FALSE) +
    
  # census units
  tmap::tm_shape(pop_2020) +
    tmap::tm_polygons(title = "Supply of park area per capita (km<sup>2</sup>)",
                      group = "Subzones: Per capita supply of park area",
                      col = "area_supplyperpop", 
                      palette = "Greens", alpha = 0.7,
                      style = "quantile",
                      border.col = "white", border.alpha = 0.5, lwd = 1) +
  tmap::tm_shape(pop_2020) +
    tmap::tm_polygons(title = "Park area per capita (km<sup>2</sup>)",
                      group = "Subzones: Per capita park area (conventional)",
                      col = "parkperpop_m2",
                      palette = "Greens", alpha = 0.7,
                      style = "quantile",
                      border.col = "white", border.alpha = 0.5, lwd = 1)
  
  
# Pipe the tmap object into tmap_leaflet() to create a leaflet widget,
# so that we can use leaflet::hideGroup().
tm %>% 
  tmap::tmap_leaflet() %>%
  leaflet::hideGroup("Buildings: Supply of park area") %>%
  leaflet::hideGroup("Subzones: Per capita supply of park area") %>% 
  leaflet::hideGroup("Subzones: Per capita park area (conventional)") 

```


```{r include = FALSE}
rm(tm, buildings_pop_sgp, parks_sgp, pop_sgp, pop_2020)
```


<br>

## Data sources

- Singapore census data from the [Department of Statistics Singapore](https://www.singstat.gov.sg/find-data/search-by-theme/population/geographic-distribution/latest-data). Released under the terms of the [Singapore Open Data Licence version 1.0](https://data.gov.sg/open-data-licence).

- Singapore subzone polygons from the [Singapore Master Plan Subzones](https://data.gov.sg/dataset/master-plan-2019-subzone-boundary-no-sea). Released under the terms of the [Singapore Open Data Licence version 1.0](https://data.gov.sg/open-data-licence).

- Singapore Master Plan Land Use Zones for the years [2014](https://data.gov.sg/dataset/master-plan-2014-land-use) and [2019](https://data.gov.sg/dataset/master-plan-2019-land-use-layer). Released under the terms of the [Singapore Open Data License](https://data.gov.sg/open-data-licence).

- Building polygons derived from map data [copyrighted](https://www.openstreetmap.org/copyright) OpenStreetMap contributors and available from https://www.openstreetmap.org. Released under the terms of the [ODbL License](https://opendatacommons.org/licenses/odbl/summary/).

- Park polygons and summarised attributes (trails, playgrounds) derived from map data [copyrighted](https://www.openstreetmap.org/copyright) OpenStreetMap contributors and available from https://www.openstreetmap.org. Released under the terms of the [ODbL License](https://opendatacommons.org/licenses/odbl/summary/).


